% file for pdflatex.exe

% continuum.tex v6.0 (c) | Copyright 2023 Daniel E. Janusch

% This file is licensed by https://raw.githubusercontent.com/drizzt536/files/main/LICENSE
% and may only be copied IN ITS ENTIRETY under penalty of law.


\documentclass[12pt]{article}
\usepackage{amssymb, latexsym, amsmath, amsthm}
\usepackage[margin=1in]{geometry}

\addtolength{\topmargin}{-0.5in}
\newtheorem{thm}{Theorem}

\renewcommand{\qedsymbol}{\blacksquare}

\begin{document}


\title{On the Continuum Hypotheses and the Correspondence of Infinite Sets' Elements with the Natural Numbers}
\author{Daniel E. Janusch\\Dedicated to Maeby, my favorite kitty}
\maketitle

\begin{section}{Introduction and Countability Theorems}\label{sec:background}
	Any ``countable" set is equinumerous to a improper subset of the natural numbers, and
	there exists a bijection between them. Any finite set has this property, the main
	question being if all \emph{infinite} sets do. In all infinite processes, one should
	think of the steps as executing simultaneously rather than separately and successively.
	$|S|$ and $C(S)$ denote the cardinality and countability of a set $S$ respectively.
	Equations (2a) and (2b) describe what will be called sum amd product ``ring operators."
	Most theorems use ``$\iff$" instead of ``$\implies$" because Theorem~\ref{thm:subsets}.1
	works as an inversion. Equation~(1a) doesn't nest groups but flattens them.
	\vspace{-1em}\\
	\begin{align}
		R\times S & :=\left\{(r,s):r\in R\land s\in S\right\}~~~~(1\text a)
		\hspace{2em}
		R\cdot S:=\left\{r\cdot s:r\in R\land s\in S\right\}~~~~~\,(1\text b)\\
		R(\vec v\,) & :=\left\{\sum_{n=1}^{\text{len}\,\vec v}(r_n\cdot\vec v_n):r_k\in R\right\}~~~(2\text a)
		\hspace{2.1em}
		R[\,\vec v\,]:=\left\{\prod_{n=1}^{\text{len}\,\vec v}(r_n+\vec v_n):r_k\in R\right\}~~(2\text b)\\
		\aleph_0 & := |\mathbb N|\\
		C(S) & :=\begin{cases}
			1,\,\left[S\text{ countable}\right]\phantom{un}~~(|S|\leqslant\aleph_0)\\
			0,\,\left[S\text{ uncountable}\right]~~(|S|>\aleph_0)\\
		\end{cases}
	\end{align}

	\begin{thm}\label{thm:finite unions}
		\emph{
			Singlular Unions of Countable Sets\\
			\indent If any sets $R$ and $S$ are countable, $R(a)\cup S(b)$ is countable
			for all numbers $a$ and $b$.\vspace{0.4em}\\
			\indent1. $C(R)\land C(S)\iff C(R\left(a\right)\cup S\left(b\right))~\forall a,b$
			\hspace{1em}
			2. $C(R)\land C(S)\iff C(R\left(a\right)\cup S[\,b\,])~\forall a,b$\\
			\indent3. $C(R)\land C(S)\iff C(R[\,a\,]\cup S\left(b\right))~\forall a,b$
			\hspace{1em}
			4. $C(R)\land C(S)\iff C(R[\,a\,]\cup S[\,b\,])~\forall a,b$
		}
	\end{thm}\begin{proof}
		$R$ and $S$ being countable implies $R(a)$ and $S(b)$ are countable as well
		because ring operators keep cardinalities the same. (case 1) Interlace the elements of
		$R(a)$ with those of $S(b)$, skipping over any duplicates. If either set is
		finite, append its elements to the beginning of the other set sans duplicates.
		Similar logic works with the other cases.
	\end{proof}

	\begin{thm}\label{thm:multiplications}
		\emph{
			Cartesian Products and Element-wise Multiplications of Countable Sets\\
			\indent If any sets $R$ and $S$ are countable, both $R\times S$ and
			$R\cdot S$ are countable.\vspace{0.4em}\\
			\indent1. $C(R)\land C(S)\iff C(R\times S)$\\
			\indent2. $C(R)\land C(S)\iff C(R\cdot S)$
		}
	\end{thm}\pagebreak\begin{proof}
		To create $R\times S$, put all possible ordered pairs into a table as shown.
		Going along the diagonals, including every element in the output, and not
		``missing" any, one can show there is a bijection between $R\times S$ and
		$\mathbb N$. The numbers and colons in the table are the output indices, the
		first 25 being labeled if in view. Similar logic works for $R\cdot S$. This
		method is also algorithmically viable (see Section~\ref{sec:algorithm code}),
		and the output indices relate strongly to triangle numbers.\vspace{-0.3em}\\
		\centerline{\begin{array}{|c|c|c|c|c|}
			\hline \phantom{0}1:(R_1, S_1) & \phantom{0}2:(R_1, S_2) & \phantom{0}4:(R_1, S_3) & \phantom{0}7:(R_1, S_4) & \cdots\\
			\hline \phantom{0}3:(R_2, S_1) & \phantom{0}5:(R_2, S_2) & \phantom{0}8:(R_2, S_3) & 12:(R_2, S_4) & \cdots\\
			\hline \phantom{0}6:(R_3, S_1) & \phantom{0}9:(R_3, S_2) & 13:(R_3, S_3) & 18:(R_3, S_4) & \cdots\\
			\hline 10:(R_4, S_1) & 14:(R_4, S_2) & 19:(R_4, S_3) & 25:(R_4, S_4) & \cdots\\
			\hline \vdots & \vdots & \vdots & \vdots & \ddots\\
			\hline
		\end{array}}\vspace{-1.1em}
	\end{proof}

	\begin{thm}\label{thm:subsets}
		\emph{
			Subsets, Intersections, and Differences of Countable Sets\\
			\indent If any set $R$ is countable, any loose subset $S$ of $R$ is also countable.
			\vspace{0.4em}\\
			\indent1. $C(R)\land S\subseteq R\implies C(S)$\\
			\indent2. $C(R)\land C(S)\implies C(R\cap S)$\\
			\indent3. $C(R)\implies C(R\smallsetminus S)$
		}
	\end{thm}\begin{proof}
		Since subsets of $\mathbb N$ are countable and $R$ is countable, all subsets of $R$
		must also be countable due to the bijection between $R$ and $\mathbb N$. This also
		implies that set intersections and differences are countable because these operations
		return subsets of the left input. Case 3 doesn't require $S$ to be countable because
		it is uncountable and $R$ is countable, then it is just more likely that the output
		will be the empty set.
	\end{proof}

	\begin{thm}\label{thm:rings}
		\emph{
			Two-Argument Ring Operators on Countable Sets\\
			\indent If any set $R$ is countable, $R(a, b)$ and $R[a, b]$ are also
			countable for any constants $a, b$.\vspace{0.4em}\\
			\indent1. $C(R)\iff C(R(a, b))~\forall a,b$\\
			\indent2. $C(R)\iff C(R[\,a, b\,])~\forall a,b$
		}
	\end{thm}\begin{proof}
		$R(a,b)$ can be tabulated with axes $n$ and $k$, where the corresponding
		table box is $aR_n + bR_k$. This is countable in the same way as
		Theorem~\ref{thm:multiplications}. Similar logic works for $R[a,b]$.\\
		\centerline{\begin{array}{|c|c|c|c|c|c|} % 6 columns
			\hline aR_1 + bR_1 & aR_1 + bR_2 & aR_1 + bR_3 & aR_1 + bR_4 & aR_1 + bR_5 & \cdots\\
			\hline aR_2 + bR_1 & aR_2 + bR_2 & aR_2 + bR_3 & aR_2 + bR_4 & aR_2 + bR_5 & \cdots\\
			\hline aR_3 + bR_1 & aR_3 + bR_2 & aR_3 + bR_3 & aR_3 + bR_4 & aR_3 + bR_5 & \cdots\\
			\hline aR_4 + bR_1 & aR_4 + bR_2 & aR_4 + bR_3 & aR_4 + bR_4 & aR_4 + bR_5 & \cdots\\
			\hline \vdots & \vdots & \vdots & \vdots & \vdots & \ddots\\
			\hline
		\end{array}}
	\end{proof}

	\begin{thm}\label{thm:powers}
		\emph{
			Finite Natural Powers of Countable Sets\\
			\indent If any set $R$ is countable, $R^n$ is also countable for all natural
			numbers $n$.\vspace{0.4em}\\
			\indent1. $C(R)\iff C(R^n)~~\forall n\in\mathbb N_0<\infty$
		}
	\end{thm}\begin{proof}
		$R^n$ can be factored as $R\times R^{n-1}$ (eqn.~1a flattens outputs), which is countable if $R^{n-1}$ is countable (Theorem~\ref{thm:multiplications}.1). This can be applied recursively until it becomes countable if $R^2$ or $R\times R$ is
		countable, which is countable via Theorem~\ref{thm:multiplications}.1. If $n=0$, the
		output set is all the groups of zero elements from $R$, or just $\emptyset$, which is countable
		since $\left|\emptyset\right|$ is finite.
	\end{proof}

	\begin{thm}\label{thm:infinite unions}
		\emph{
			Countably Infinite Unions of Countably Infinite Sets\\
			\indent If all sets $R_i$ are countable, their union is also countable.
			\vspace{0.4em}\\
			\indent1. $C(R_i)\forall i\iff C\!\left(\displaystyle\bigcup_{i=1}^{\aleph_0}R_i
			\right)$
		}
	\end{thm}\begin{proof}
		To create the union, give each set $R_i$ a column in a table and put all its elements
		sequentially in that column going down. Do this for every $R_i$, and using the same
		method as in Theorem~\ref{thm:multiplications}, one can show that this set bijects
		the naturals. Every column represents a different $R_i$ and each row represents a
		different $\left(R_i\right)_j$ or $R_{i,j}$. The numbers before the colons are the
		output index, the first 85 shown if in view. This is algorithmically viable
		(see Section~\ref{sec:algorithm code}).\vspace{0.5em}
		\centerline{\begin{array}{|c|c|c|c|c|c|c|c|} % 8 columns
			\hline \phantom{0}1:R_{1,1} & \phantom{0}2:R_{2,1} & \phantom{0}6:R_{3,1} & \phantom{0}7:R_{4,1} & 15:R_{5,1} & 16:R_{6,1} & 28:R_{7,1} & \cdots\\
			\hline \phantom{0}3:R_{1,2} & \phantom{0}5:R_{2,2} & \phantom{0}8:R_{3,2} & 14:R_{4,2} & 17:R_{5,2} & 27:R_{6,2} & 30:R_{7,2} & \cdots\\
			\hline \phantom{0}4:R_{1,3} & \phantom{0}9:R_{2,3} & 13:R_{3,3} & 18:R_{4,3} & 26:R_{5,3} & 31:R_{6,3} & 43:R_{7,3} & \cdots\\
			\hline 10:R_{1,4} & 12:R_{2,4} & 19:R_{3,4} & 25:R_{4,4} & 32:R_{5,4} & 42:R_{6,4} & 49:R_{7,4} & \cdots\\
			\hline 11:R_{1,5} & 20:R_{2,5} & 24:R_{3,5} & 33:R_{4,5} & 41:R_{5,5} & 50:R_{6,5} & 62:R_{7,5} & \cdots\\
			\hline 21:R_{1,6} & 23:R_{2,6} & 34:R_{3,6} & 40:R_{4,6} & 51:R_{5,6} & 61:R_{6,6} & 72:R_{7,6} & \cdots\\
			\hline 22:R_{1,7} & 35:R_{2,7} & 39:R_{3,7} & 52:R_{4,7} & 60:R_{5,7} & 73:R_{6,7} & 85:R_{7,7} & \cdots\\
			\hline \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \ddots\\
			\hline
		\end{array}}
	\end{proof}
\end{section}

\begin{section}{Countabilities of Common Sets}\label{sec:applications}
	\begin{subsection}{Countability of the Integers}\label{subsec:applications.integers}
		Claim: $\mathbb Z$ is a countable set; $C(\mathbb Z)\equiv1$.
		\begin{proof}
			$R:=\mathbb N_0$, $S:=\mathbb N_1$. $R$ and $S$ are countable axiomatically
			(axiom of extensionality), being the naturals themselves. $\mathbb Z\equiv
			R(1)\cup S(-1)$; this is countable via Theorem~\ref{thm:finite unions}.1.
		\end{proof}
	\end{subsection}

	\begin{subsection}{Countability of the Rationals}\label{subsec:applications.rationals}
		Claim: $\mathbb Q$ is a countable set; $C(\mathbb Q)\equiv1$.
		\begin{proof}
			$\mathbb Q$ has the same countability as $\mathbb Z\times \mathbb N_1$.
			$C(\mathbb Z)\equiv1$ via Section~\ref{subsec:applications.integers} and
			$C(\mathbb N_1)\equiv1$ axiomatically. Thus $C(\mathbb Z\times \mathbb N_1)\equiv1$
			via Theorem~\ref{thm:multiplications}.1. This and Theorem~\ref{thm:multiplications} derive
			from Cantor's enumeration of countable collections of countable sets.
		\end{proof}
	\end{subsection}

	\begin{subsection}{Countability of the Reals from Zero to One}\label{subsec:applications.reals 0 to 1}
		Claim: $\{x\in\mathbb R:0\leqslant x<1\}$ is a countable set;
		$C(\mathbb R_{0\leqslant x<1})\equiv1$.
		\begin{proof}
			Let $R$ be a set where $\forall i\in\mathbb N_0$, $R_i := $ (``0." $+$ reversed digits
			of $i$), ie: $R_{2760}=0.0672$ (see Section~\ref{sec:further algorithms} for formula). $R$
			is countable by definition, and contains every possible sequence of digits in $[0,1)$. The
			sequence trends towards different values with different sub-series;
			Section~\ref{subsec:applications.reals} explains this further. $R$ is used in the next proof.
		\end{proof}
	\end{subsection}

	\pagebreak\begin{subsection}{Countability of the Non-Negative Reals}
	\label{subsec:applications.non negative reals}
		Claim: $\mathbb R_{\geqslant0}$ or equivalently $\left\{x\in\mathbb R:x\geqslant0\right\}$,
		is a countable set; $C(\mathbb R_{\geqslant0})\equiv1$.
		\begin{proof}
			$C(\mathbb R_{0\leqslant x<1})\equiv1$ via Section~\ref{subsec:applications.reals 0 to 1}.
			$\mathbb R_{\geqslant0}=\displaystyle\bigcup_{i=1}^{\aleph_0}S_i$ where
			$S_i:=R[i-1]$. This is countable via Theorem~\ref{thm:infinite unions}; the following
			table illustrates this. $\forall x\in\mathbb R_{\geqslant0}$,
			there exists a sequence of $s_i\in S_{\lceil x\rceil}$, with increasing indices, where
			$\displaystyle\lim_{i\in\mathbb N\to\infty}s_i=x$. This is because $S_n$ contains every
			sequence of digits in $[n-1,n)$.
			$S$ is used in the next proof.\vspace{0.5em}\\
			\centerline{\begin{array}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|} % 13 columns
				\hline 0 & 0.1 & 0.2 & 0.3 & 0.4 & 0.5 & 0.6 & 0.7 & 0.8 & 0.9 & 0.01 & 0.11 & 0.21 & \cdots\\
				\hline 1 & 1.1 & 1.2 & 1.3 & 1.4 & 1.5 & 1.6 & 1.7 & 1.8 & 1.9 & 1.01 & 1.11 & 0.21 & \cdots\\
				\hline 2 & 2.1 & 2.2 & 2.3 & 2.4 & 2.5 & 2.6 & 2.7 & 2.8 & 2.9 & 2.01 & 2.11 & 0.21 & \cdots\\
				\hline 3 & 3.1 & 3.2 & 3.3 & 3.4 & 3.5 & 3.6 & 3.7 & 3.8 & 3.9 & 3.01 & 3.11 & 0.21 & \cdots\\
				\hline 4 & 4.1 & 4.2 & 4.3 & 4.4 & 4.5 & 4.6 & 4.7 & 4.8 & 4.9 & 4.01 & 4.11 & 0.21 & \cdots\\
				\hline \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \ddots\\
				\hline
			\end{array}}
			
		\end{proof}
	\end{subsection}

	\begin{subsection}{Countability of the Reals}\label{subsec:applications.reals}
		Claim: $\mathbb R$ is a countable set which implies $2^{\aleph_0}=\aleph_0$.
		\begin{proof}
			$\mathbb R_{\geqslant0}$ is countable via Section~\ref{subsec:applications.non negative reals}.
			$\mathbb R_{\geqslant0}(1, -1)\equiv\mathbb R$. This is countable via
			Theorem~\ref{thm:rings}.1. This conclusion can be further reinforced by
			the algorithmic methods in Section~\ref{sec:algorithm code}.
		\end{proof}
	\end{subsection}

	\begin{subsection}{Countabilities of Miscellaneous Number Classes}
	\label{subsec:applications.miscellaneous}
		\begin{subsubsection}{Algebraic and Transcendental Numbers}
			\label{subsubsec:applications.misc.reals}
			Claim: $\mathbb A$ and $\mathbb T$ are countable sets over both $\mathbb R$ and $\mathbb C$.
			\begin{proof}
				$\mathbb R$ is a countable set via Section~\ref{subsec:applications.reals}.
				Since the algebraic reals and transcendental reals are both subsets of the
				reals, they are countable via Theorem~\ref{thm:subsets}.1. They are also
				countable over the complex numbers using Section~\ref{subsec:applications.complex}
				and the same logic.
			\end{proof}
		\end{subsubsection}

		\begin{subsubsection}{Imaginary Numbers}\label{subsubsec:applications.misc.imaginary}
			Claim: $\mathbb I$ is a countable set; $C(\mathbb I)\equiv1$.
			\begin{proof}
				$\mathbb R$ is a countable set via Section~\ref{subsec:applications.reals}.
				$\mathbb R(\sqrt{-1})\cup\emptyset(1)\equiv\mathbb I$. This is countable via
				Theorem~\ref{thm:finite unions}.1 becuase the union of any set and the null set
				is itself.
			\end{proof}
		\end{subsubsection}
	\end{subsection}

	\begin{subsection}{Countability of the Complex Numbers}\label{subsec:applications.complex}
		Claim: $\mathbb C$ is a countable set; $C(\mathbb C)\equiv1$.
		\begin{proof}
			$\mathbb R$ and $\mathbb I$ are countable sets via Sections \ref{subsec:applications.reals}
			and \ref{subsubsec:applications.misc.imaginary} respectively. $\mathbb R(1, \sqrt{-1})\equiv\mathbb C$
			is countable via Theorem~\ref{thm:rings}.1. $\mathbb R\times\mathbb I=\mathbb C$ is
			countable via Theorem~\ref{thm:multiplications}.1.
		\end{proof}
	\end{subsection}
\end{section}

\begin{section}{Generalized Continuum Hypothesis}\label{sec:GCH}
	Claim: GCH has the same truth value as CH, and powersets don't change sizes of infinity.
	\begin{proof}
		Let $S$ be a countably infinite set; it cannot be finite or this proof does not work.
		$P :=\mathcal P(S)$. Let $R_k$ denote the set of elements in $P$ with $k$ sub-elements
		($S^k$); these are countable via Theorem~\ref{thm:powers}. Observe that $P$ has
		$\aleph_0^0$ elements with 0 sub-elements, $\aleph_0^1$ elements with 1 sub-element, et
		cetera, and generally $\aleph_0^k$ elements with $k$ sub-elements, for
		$k\leqslant \aleph_0$. Imagine a $k$-dimensioned table of integers and listing them in
		a $k$-dimensional spiral, or recursively combining two table dimensions into one using
		the Theorem~\ref{thm:multiplications} table method until the table gets to one dimension.
		$\displaystyle P=\bigcup_{i=1}^{\aleph_0}R_{i-1}$ is countable via
		Theorem~\ref{thm:infinite unions}. This contradicts Cantor's Diagonal Argument (see
		Section~\ref{sec:diagonal argument}) and implies that uncountable sets cannot be created
		through powersets of countable sets, and either cannot be created at all, or need a much
		stronger method to create them.
	\end{proof}
\end{section}

\begin{section}{Cantor's Diagonal Argument}\label{sec:diagonal argument}
	According to Georg Cantor in 1891$^\text{[ref \ref{ref:diagonal argument},
	\ref{ref:cantor-1891}]}$, If someone is trying to list all the real numbers, they can
	always find a number that is not in the list using his ``Diagonal Argument" which
	accomplishes the same as the following, though for reals instead of
	naturals: suppose someone is trying to make a set $S$ with every natural number
	($\mathbb N_1$). They could first skip 1 and add 2 to $S$. Then they add 3, then 4, 5, 6,
	et cetera forever. No matter how many natural numbers they add, they can always find one
	not in it; $\max(S)+1$. This seems to be implying that there is not a bijection between the
	natural numbers and themselves since in the limit towards infinity time, they've used up all
	infinity natural numbers indexing $S$ and yet not all of them are in $S$. Well of course, because
	they \emph{skipped} 1. Due to this possibility of skipping elements, the diagonal argument doesn't always accurately depict
	the countability of sets, and it can just as easily skip elements in the set of real numbers.
\end{section}

\begin{section}{Enumerating the Continuum Algorithmically}\label{sec:algorithm code}
	The following NodeJS code prints out real numbers to stdout up to the maximum allowed big integer
	delimited by a comma-space pair, the problem being that it prints out both $\pm0$, though this
	is easily resolvable. The more verbatim version as well as the C version of the source code are at
	References~\ref{ref:js-code}~and~\ref{ref:c-code}. The complexity of finding the $n^\text{th}$ real
	number with this method is around $\Theta(\log n)$, assuming basic operations such as bit shifts,
	comparisons, addition, and multiplication run in $\mathcal O(1)$. isqrt($n$) is less than
	$\mathcal O(\log_2 n)$, and contains most of the complexity of the formula. The code didn't start
	to slow down noticably until around $10^{100}$, and at $10^{10,000}$ only took 2.4 seconds (rather
	than 0.2ms at the start).

	\noindent\texttt{\\
		const write = globalThis.toString().slice(8, -1).toLowerCase() === "global" ?\\
		$\indent$s => process.stdout.write.call(process.stdout, s) : // NodeJS\\
		$\indent$console.log; // Browser\\
		\\
		function isqrt(n) \{ // floored square root\\
		$\indent$if (n < 2n) return n;\\
		\\
		$\indent$var cur = n >> 1n, prev; // current, previous\\
		\\
		$\indent$do [cur, prev] = [cur + n / cur >> 1n, cur];\\
		$\indent$while (cur < prev);\\
		\\
		$\indent$return prev;\\
		\}\\
		\\
		for (var i = 0n ;;) \{\\
		$\indent$const k = 1n + isqrt(1n + (i << 2n)) >> 1n // T(k-1) < i/2 $\leq$ T(k)\\
		$\indent$$\indent$, t = i + k*(1n - k) >> 1n; // i/2 - T(k-1)\\
		\\
		$\indent$write( // coordinates: (k, k-t-1)\\
		$\indent$$\indent$\textasciigrave\$\{i++ \% 2n ?\;"-"\;:\;""\}\$\{t\}.\textasciigrave + // integer part\\
		$\indent$$\indent$\textasciigrave\$\{k - t - 1n\}\textasciigrave.split("").reverse().join("") // fractional part\\
		$\indent$)\\
		\}
	}
\end{section}

\begin{section}{Further Algorithms}\label{sec:further algorithms}
	The axioms of Zermelo-Fraenkel Set theory required to prove the algorithmic viability of
	continuum enumeration are as follows: Axiom Schema of Specification, Axiom of Union, Axiom
	of Infinity, and Axiom of Choice (for creating sets of ordered \textit{pairs}, which is not
	the controversial part). Without the Axiom of Choice, Theorem~\ref{thm:multiplications}
	doesn't work, although the algorithm from Section~\ref{sec:algorithm code} is separate from
	set theory and assumes no axioms. The previous algorithm uses string manipulation to reverse
	the numbers and find their lengths, but the following formulas for length() and reverse() do
	the same thing mathematically.
	\begin{align}
		\text{length}\,n & := 1+\lfloor\log_b|n|\rfloor=\lceil\log_b(|n|+1)\rceil\ni\ln0\leqslant0\\
		\text{reverse}\,n & :=\text{sgn}(n)\!\!\sum_{k=0}^{\text{length}\,n}\left(\left\lfloor
		\frac{|n|}{b^k}\right\rfloor\text{mod}\,b\right)\!b^{\text{length}\,n-k-1}\\
		R_{i,j} & := s\cdot\left(i+\frac{\text{reverse}\,j}{b^{\text{length}\,j}}\right)
	\end{align}
	Where $i$ is the integer part, $s$ is the sign, $j$ is the reversed decimal part, and $b$
	is the base. $i$ and $j$ can be swapped for a different enumeration (and inverse) (see
	Reference~\ref{ref:desmos real-enum}). A formula for the greatest triangle number
	$t=T(k)\leqslant n$, where $n,k\in\mathbb N_0$ (used in Section~\ref{sec:algorithm code}):\\
	\begin{align}
		T(n) := \frac{n^2+n}2 & = \sum_{j=1}^nj\\
		t = T\left(\left\lceil\frac{\left\lfloor\sqrt{8n+1}\right\rfloor}2\right\rceil-1\right)
		& = T\left(\left\lfloor\dfrac{\sqrt{8n+1}}2\right\rceil-1\right)
	\end{align}
	Because of the bijection, there is also an inverse of the Section~\ref{sec:algorithm code}
	algorithm. The simplicity and symmetry of the inverse is striking due to the complexity
	of the original function.\\\\
	\noindent\texttt{function inverse(string) \{\\
		$\indent$const match = /\textasciicircum-??($\backslash$d+)$\backslash$.($\backslash$d+)\$/.exec(string)\\
		$\indent\indent$, x = BigInt(match[1]) // integer part\\
		$\indent\indent$, y = BigInt(match[3].split("").reverse().join("")); // decimal part\\
		$\indent$return (x+y)**2n + 3n*x + y + BigInt(string[0] === "-");\\
		\}
	}
\end{section}

\begin{section}{Second Enumeration}\label{sec:second enumeration}
	There are more enumerations than just what was previously mentioned. A (mostly unrigorous) proof
	will be presented for the second of the following equations, both of which derive from a similar
	idea. Something similar works for any natural number base
	(Reference~\ref{ref:python second enum family}: Python code,
	Reference~\ref{ref:desmos second enum family}: desmos graph), instead of just 2, although it works
	best with prime bases.
	\begin{align}
	\lim_{n,k\in\mathbb N_1\to\aleph_0}\left\{k\!\left(\dfrac{2i+1}{2^{\lfloor\log_2i\rfloor}}-3\right):i\in\mathbb N_1<2^n\right\}=\mathbb R\\
	\bigcup_{n=1}^{\aleph_0}\dfrac{2n+1}{2^{1+\lfloor\log_2n\rfloor}}\equiv\bigcup_{n=1}^{\aleph_0}\dfrac{2n+1}{2^{\text{length}_2(n)}}=\mathbb R_{(1,2)}
	\end{align}
	\begin{proof} This proof will first prove that $S_{\aleph_0}$ is equivalent to $\mathbb R_{[0,1)}$,
	and then proves that it must be countable.
		\begin{align*}
			& S_0:=\{0\}\\
			& S_n:=S_{n-1}\cup\left\{x+\dfrac1{2^n}:x\in S_{n-1}\right\}=\dfrac{\{k\in\mathbb N:0\le k<2^n\}}{2^n}\\
			& S_{\aleph_0}:=\lim_{n\in\mathbb N\to\aleph_0}S_n=\mathbb R_{[0,1)}
		\end{align*}
		$S_{\aleph_0}$ must equal [0,1) because 0 and 1 are its bounds and it has the same cardinality
		as the real numbers. It starts with 0 and only adds to it conditionally so its elements never
		go below 0, and $\displaystyle\sum_{n=1}^\infty\dfrac1{2^n}=1$, meaning it can't extend past 1.
		$|S_n|=2^n$ which implies $|S_{\aleph_0}|=2^{\aleph_0}$, the same cardinality as
		$|\mathcal P(\mathbb N)|$ or $|\mathbb R|$. $S_{\aleph_0}$ is countable via Cantor's Enumeration
		of Countable Collections of Countable sets, since there are countably infinitely many countable
		processess involved in constructing $S_{\aleph_0}$. The ``countable collection" refering to the
		indices ``$n$," and the ``processes" being the evaluation of $S_n$ given $S_{n-1}$, each of which
		processes is itself finite, and thus countable. The same results arise when starting with $S_0=\{1\}$
		and subtracting subsequent inverse powers of 2 rather than adding.
	\end{proof}
\end{section}

\begin{section}{Conclusion}\label{sec:conclusion}
	Since there exists a bijection between any infinite set and its powerset (GCH), there is no
	set with a cardinality strictly or loosely in between the naturals and reals (CH), because
	they are the same cardinality. This also implies that $\aleph_n=\beth_n=\beth_0$ for all
	natural numbers $n$, which makes sense intuitively because $2^\infty=\infty$. Also if any
	sequence of natural numbers is concatenated, it will always create a new natural number,
	meaning every element in $\mathcal P(\mathbb N_0)$ corresponds to an element in
	$\mathbb N_0$. This all could have disasterous consequences for set theory because CH is
	provably undecidable in some models and yet provably decidable. This means either
	Zermelo-Fraenkel Set Theory with the axiom of choice (ZHC) is unsound, the aforementioned
	models are invalid, or all of the decidability proofs are invalid. The entire foundation of
	aleph numbers, beth numbers, and sizes of infinities could be entirely flawed. CH implies
	the Gimel Hypothesis is true according to Reference~\ref{ref:gimel}, and implies Wetzel's
	problem is false according to Reference~\ref{ref:wetzel}. A truth value for CH or GCH is
	not asserted here bacause the specifics of CH are unclear, and they are important for its
	truth value. If the reals have to be bigger than the naturals, it is false, but if it just
	requires that there is no set in between them, it is true.
\end{section}

\pagebreak\begin{section}{References}\label{sec:references}
	\begin{enumerate}
		\item\url{https://en.wikipedia.org/wiki/Cantor's\_diagonal\_argument}\\
		\label{ref:diagonal argument}
		\text{Wikipedia page with background for Section~\ref{sec:diagonal argument}}

		\item\url{https://www.digizeitschriften.de/dms/img/?PID=GDZPPN002113910&physid=phys84#navi}\\
		\label{ref:cantor-1891}
		\text{Georg Cantor's 1891 article with the diagonal argument. Same source as on wikipedia.}

		\item\url{https://en.wikipedia.org/wiki/Continuum\_hypothesis}\\
		\label{ref:continuum}
		\text{Wikipedia page with elaboration on Section~\ref{subsec:applications.reals}}

		\item\url{https://www.github.com/drizzt536/files/tree/main/TeX/continuum}\\
		\label{ref:files}
		\text{The files for the most recent public version of this pdf and the \LaTeX~code}

		\item\url{https://raw.githubusercontent.com/drizzt536/files/main/JavaScript/continuum.js}\\
		\label{ref:js-code}
		\text{The raw JavaScript source code for the Section~\ref{sec:algorithm code}}

		\item\url{https://raw.githubusercontent.com/drizzt536/files/main/C/continuum.c}\\
		\label{ref:c-code}
		\text{The raw C source code for the Section~\ref{sec:algorithm code}}

		\item\url{https://www.desmos.com/calculator/3pmv7lehhm}\\
		\label{ref:desmos real-enum}
		\text{Desmos with real number enumeration 1 formulas for Section~\ref{sec:further algorithms}}

		\item\url{https://github.com/drizzt536/files/blob/main/TeX/continuum/real-enum-formula/formula.pdf}\\
		\label{ref:pdf real-enum}
		\text{pdf with a fully-expanded real number enumeration 1 formula for Section~\ref{sec:further algorithms}}

		\item\url{https://github.com/drizzt536/files/blob/main/Python/continuum.py}\\
		\label{ref:python second enum family}
		\text{second enumeration family python code}

		\item\url{https://www.desmos.com/calculator/dbcdgkinga}\\
		\label{ref:desmos second enum family}
		\text{second enumeration family desmos graph}

		\item\url{https://en.wikipedia.org/wiki/Gimel\_function\#The\_gimel\_hypothesis}\\
		\label{ref:gimel}
		\text{Gimel Hypothesis Wikipedia page}

		\item\url{https://en.wikipedia.org/wiki/Wetzel's\_problem}\\
		\label{ref:wetzel}
		\text{Wikipedia page for Wetzel's Problem}
	\end{enumerate}\\
	Editors:
	\begin{itemize}
		\item Daniel E. Janusch
		\item Valerie Janusch, my mom
	\end{itemize}
	\vspace{3em}\\
	This document is licensed under https://raw.githubusercontent.com/drizzt536/files/main/LICENSE
	and may only be copied IN ITS ENTIRETY under penalty of law.
\end{section}

\end{document}
